Improving Software Maintainability through Automated Refactoring of Code Clones

### 2019-10-26 saner/paper.tex
Updates from Overleaf

%Clone detection tools often use thresholds to determine which clones should be considered. Most clone detection tools use the number of lines \cite{kamiya2002ccfinder, svajlenko2016bigcloneeval}, statements \cite{higo2013revisiting, kamalpriya2017enhancing} and/or tokens \cite{ragkhitwetsagul2019siamese, jiang2007deckard} as thresholds to determine whether clones should be considered. We outline the shortcomings of these thresholds to find clones such that the maintainability of the source code is improved when refactored, and propose new thresholds to control this characteristic.
% Current clone detection techniques are limited to measuring the size of clones to determine whether they should be considered. Because of this, the output of clone detection tools is often of limited assistance to the developer in the refactoring process.
To reduce duplication of a codebase, there are several refactoring techniques that can be applied. The most used refactoring technique to deal with code clones is ``Extract Method'' \cite{fowler2018refactoring}. This technique entails that the duplicated code is moved to a new method, which is then called from all locations where the duplicated code was used. To be able to execute such a refactoring, several preconditions apply \cite{tsantalis2015assessing}.

In this study, we analyze which preconditions apply to determine whether a clone can be automatically refactored. We list criteria that influence the maintainability of clones when refactored. We evaluate the context of clones to determine which refactoring techniques are best suited. We define the context of clones as the inheritance relation between clone instances in a clone class and the location of clone instances in the codebase. We showcase the proposed technique as a tool that automatically refactors clones by extracting new methods out of duplicated code.

@@

\section{Background}\label{sec:background}
In this section, we describe relevant code clone and clone refactoring research.

@@

\subsubsection{Refactorability} \label{sec:refactorability}
Several factors may obstruct the possibility to extract code to a new method~\cite{tsantalis2015assessing}:
\begin{itemize}
    \item \textbf{Complex Control Flow}: This clone contains \texttt{break}, \texttt{continue} or \texttt{return} statements, obstructing the possibility of method extraction.
    \item \textbf{Spans Part Of A Block}: This clone spans a part of a block statement. An example of this is shown in Fig.~\ref{fig:refactorabilitypartialblock}.
    \item \textbf{Is Not A Partial Method}: If the clone does not fall in the ``Partial method'' category of Sec.~\ref{sec:location}, the ``Extract Method'' refactoring technique cannot be applied.
    \item \textbf{Returns Multiple Values}: If a clone modifies or declares multiple pieces of data that it should return.
    \item \textbf{Top-Level Non-Statement}: If one of the top-level AST nodes of the clone is not a statement. For instance, if a (part of) an anonymous class is cloned.
    \item \textbf{Can Be Extracted}: This clone is a fragment of code that can directly be extracted to a new method. Then, based on the relation between the clone instances, further refactoring techniques can be used to refactor the extracted methods (for instance ``pull up method'' for clones in sibling classes).
\end{itemize}
Clones that do not fall in the \textit{Can Be Extracted} category may require additional transformations or other techniques to refactor. CR only refactors the clones that \textit{Can Be Extracted}.

\subsubsection{AST Transformation} % TODO: Big todo --> Find literature, introduce proper reasoning why I need to do this.
rtCR uses JavaParser~\cite{smith20e17javaparser}: an AST-parsing library that allows to modify the AST and write it back to source code. To refactor clones, CR creates a new method declaration and moves all statements from a clone instance in the clone class to the new method. This method is placed according to the relation between the clone instances (see Sec.~\ref{sec:relation}). CR analyzes the source code of the extracted method and populates it with the following properties:
\begin{itemize}
  \item \textbf{Parameters}: For each variable used that is not accessible from the scope of the extracted method.
  \item \textbf{A return value}: If the method modifies or declares local data that is needed outside of its scope, or if the cloned fragments already returned data.
  \item \textbf{Thrown exception}: If the method throws an uncaught exception that is not a \texttt{RuntimeException}.
\end{itemize}
CR then removes all cloned code and replaces it with a method call to the newly created method. In case of a return value, CR either assigns the call result, declares it or returns it accordingly.

\subsubsection{Characteristics of the extracted method}\label{sec:characteristics}
We define the following characteristics of the extracted method and/or the call:
\begin{itemize}
\item \textbf{Tokens}: The number of tokens in the body of the method.
\item \textbf{Relation}: The relation category (Sec.~\ref{sec:relation}) by which this methods' location was determined.
\item \textbf{Returns}: Whether the method calls return, declare, assign or don't use any data from the extracted method.
\item \textbf{Parameters}: The number of parameters the extracted method has.
\end{itemize}
We hypothesize that these characteristics are the main factors influencing the impact on the maintainability of the system as a result of refactoring the clone.

\subsubsection{Impact on maintainability metrics} \label{sec:metrics}
CR measures the impact on maintainability metrics of the refactored source code for each clone class that is refactored. These metrics are derived from Heitlager et al.~\cite{heitlager2007practical}. This paper defines a set of metrics to measure the maintainability of a system. For each of these metrics, risk profiles are proposed to determine the maintainability impact on the system of a whole.

To determine whether the maintainability improves when refactoring a single clone, we need to measure the impact of small-grained changes. Because of that, we measure only a subset of the metrics~\cite{heitlager2007practical} and focus on the absolute metric changes (instead of the risk profiles). The subset of metrics we decided to focus on are all metrics that are measured on method level (as the other metrics show a lesser impact on the maintainability for these small changes). These metrics are:
\begin{itemize}
\item \textbf{Duplication}: In Heitlager et al.~\cite{heitlager2007practical} this metric is measured by taking the amount of duplicated lines. We decided to use the amount of duplicated tokens part of a clone class instead, to have a stronger reflection of the impact of the refactoring by measuring a more fine-grained system property.
%\item \textbf{Volume}: The more code a system has, the more code has to be maintained. The paper~\cite{heitlager2007practical} measures volume as lines of code. As with duplication, we use the number of tokens instead.
\item \textbf{Method Complexity}: Heitlager et al.~\cite{heitlager2007practical} use MCCabe complexity~\cite{mccabe1976complexity} to calculate their complexity metric. The MCCabe complexity is a quantitative measure of the number of linearly independent paths through a method.
\item \textbf{Method Interface Size}: The number of parameters that a method has. If a method has many parameters, the code may become harder to understand and it is an indication that data is not encapsulated adequately in objects~\cite{fowler2018refactoring}.
\item \textbf{Method size}: The longer a method is, the harder it becomes to comprehend and maintain \cite{heitlager2007practical}. The ``extract method'' refactoring technique is often used to split up long methods \cite{charalampidou2018structural}. When refactoring duplicated code fragments in the body of a method, this could improve maintainability of the methods from which the duplication was removed because of their reduced size.
\end{itemize}

\section{Experimental setup} \label{sec:experimentalsetup}
\todo{This doesn't say anything: explain how the section is organized}
In this section, we describe the setup of our experiments.

\subsection{Corpus}
We ran all our experiments using CR on a corpus of open source Java projects. This corpus is derived from the corpus of a study that uses machine learning to determine the suitability of Java projects on GitHub for data analysis purposes~\cite{githubCorpus2013}.

CR requires all libraries of the projects it analyses, to find the full paths of all referenced symbols in the source code (see Sec.~\ref{sec:t1r}). Because of that requirement, we decided to filter the corpus to only projects using the Maven build automation system. We created a set of scripts to prepare such a corpus with all dependencies included.

This procedure results in 2.267 Java projects including all their dependencies. The projects vary in size and quality. The total size of all projects is 14.2M lines (11.3M when excluding whitespace) over a total of 100K Java files. This is an average of 6.3K lines over 44 files per project. The largest project in the corpus is \textit{VisAD} with 502K lines.

\subsection{Tool Evaluation}
We assess the correctness of CR through unit tests and empirical validation. First, we create a set of 57 control projects to verify the correctness in many (edge) cases. These projects contain clones for each identified relation, location, and refactorability category to see whether they get correctly identified. Next, we run the tool over the corpus and manually verify samples of the acquired results. This way, we check both the correctness of the identified clones, their context, and their proposed refactoring.

We also test the correctness of the resulting code after refactoring. For this, we use a GitHub project named JFreeChart. JFreeChart has a high test coverage and working tests, which allows us to test the correctness of the program after running CR.

\subsection{Minimum clone size}
In this study, we want to find out what thresholds to use to improve maintainability if clones by those thresholds are refactored. However, when clones are very small, they may never be able to improve maintainability. The detrimental effect of the added volume of the newly created method exceeds the positive effect of removing duplication. Because of that, we perform all our experiments with a minimum clone size of 10 tokens, because smaller clones are very unlikely to improve maintainability when refactored.

\subsection{Thresholds}
Most clone detection tools can be configured using thresholds. These thresholds indicate the minimum number of lines, tokens and/or statements that must be spanned for duplicate fragments to be considered clones. Often, such thresholds are intuitively chosen~\cite{li2006cp, roy2009mutation} or based on a quartile distribution of empirical data~\cite{alves2010deriving}. Using the maintainability score we can find support for which thresholds should be chosen to increase the chance to find clones that improve maintainability when refactored.

@@

\subsection{Calculating a maintainability score}\label{sec:metricformula}
In this study, we use four metrics to determine maintainability (see Sec.~\ref{sec:metrics}). For each of these metrics, we determine their value before and after each refactored clone class, resulting in a delta metric score. For our experiments, we aggregate the deltas obtained for these metrics to draw a conclusion about the maintainability increase or decrease after applying a refactoring. We base our aggregation on the following assumptions:
\begin{itemize}
  \item All metrics are equal in terms of weight towards system maintenance effort.
  \item Higher values for the metrics imply lower maintainability.
  \item Normalizing each obtained metric delta over all deltas obtained for that metric in our dataset results in equally weighted scores.
\end{itemize}
We derived these assumptions from supporting evidence shown by Heitlager et al~\cite{heitlager2007practical} and Alves et al.~\cite{alves2010deriving}. Using the resulting aggregated maintainability score, we can argue for each refactoring whether it increases or decreases the maintainability of the system.

We normalize each obtained metric delta using the ``Standard score'', which is calculated as follows:
\begin{equation}\label{eq:scoredev}
N_{metric} = \frac {\Delta X-\mu}{\sigma}
\end{equation}
Where $\Delta X$ is a metric delta, $\mu$ is the mean of all deltas for this metric and $\sigma$ is the standard deviation of all deltas for this metric. This method works well for normalization of our data because as we divide by the standard deviation, outliers do not influence the resulting scores negatively.

We then calculate the maintainability for a specific refactoring as follows:
\begin{equation}\label{eq:scoreref}
N_{duplication} + N_{complexity} + N_{volume} + N_{parameters}
\end{equation}

@@

\section{Results} \label{sec:results}
In this section, we share the results of our experiments.

@@

\subsubsection{Relation}
Table~\ref{tab:relation} displays the number of clone classes found for the entire corpus for the identified relations between clone instances (see Sec.~\ref{sec:relation}).

@@

\subsubsection{Location}
Table~\ref{tab:location} displays the number of clone classes found for the entire corpus for different locations (see Sec.~\ref{sec:location}).

@@

From these results, we see that 74\% of clones span part of a method body (77\% if we include constructors). 8\% of clones span several methods. 6\% of clones span only global variables. Only 4\% of clones span a full declaration (method, class, constructor, etc.).

@@

These results indicate that, given our refactorability criteria, 28\% of clones can be automatically refactored. Clones in other categories may require other refactoring techniques or further transformations to be automatically refactorable.

\subsection{Thresholds}
In our corpus, CR has refactored 12.710 clone classes and measured the change in indicated metrics (see Sec.~\ref{sec:metrics}). Using the presented formulas (see Sec.~\ref{sec:metricformula}) we determine how the characteristics of the extracted method (see Sec.~\ref{sec:characteristics}) influence the maintainability of the resulting codebase after refactoring. In this section, we explore the data received by comparing the before- and after snapshots of the system for each separate refactoring. Using this data, we find supporting evidence regarding which thresholds are most likely to find clones that should be refactored to improve maintainability.

\subsubsection{Clone Token Volume}
Figure \ref{fig:maintainabilityscore} shows the obtained results when plotting the clone volume vs the average delta maintainability score. We define the \textit{token volume} as the combined number of tokens in all clone instances in a refactored clone class. For higher token volume numbers we have fewer refactorings that refactor such clones, because of which we represent the x-axis as a logarithmic scale. The trendline intersects the ``zero'' line (maintainability does not increase nor decrease) at a token volume of 63.

@@

\subsubsection{Relation}
Table~\ref{tab:relation_refactor} shows our data regarding how different relations influence maintainability. We have marked rows based on less than 100 refactorings red, as their result does not have statistical significance.

The displayed relations are ordered by their scores. Overall, the scores do not deviate much (-0.15 to 0.23), indicating that the relation between clones has a minor impact on maintainability. Overall, we see that common hierarchy clones have the highest maintainability, whereas unrelated clones have the lowest maintainability.

\subsubsection{Return Value}
Table~\ref{tab:return} shows how the return value of the extracted method influences the maintainability of the resulting system.

@@

The displayed return categories are ordered by their scores. Overall, the scores do not deviate much (-0.18 to 0.19), indicating that the return category has a minor impact on maintainability. When the result of the call to the extracted method is directly returned, the maintainability score is the highest. When no value is returned, the maintainability score is the lowest. The main reason that no return value ends up lowest, is that it is linked to a higher number of parameters required for the extracted method.

@@

\subsection{Clone Context}
Regarding clone context, our results indicate that most clones (37\%) are in a common class. This is favorable for refactoring because the extracted method does not have to be moved after extraction. 24\% of clones are in a common hierarchy. These refactorings are also often favorable. Another 24\% of clones are unrelated, which is often unfavorable because it often requires a more comprehensive refactoring. 15\% of clones are in an interface.

@@

\subsection{Extract Method}
28\% of clones can be refactored using the ``Extract Method'' refactoring technique (50\% if we limit our searching scope to method bodies). About 25\% of clones do not span part of a method, because of which they cannot be refactored. Many clones (23\%) do not have a statement as top-level AST-Node. Upon manual inspection, we noticed that the main reason for this is when clones are found in lambda functions or anonymous classes. About 15\% of clones span only part of an AST-Node.

\subsection{Refactoring}
In Fig.~\ref{fig:maintainabilityscore} we see an increase in maintainability for refactoring larger clone classes. The tipping point, between a better maintainable refactoring and a worse maintainable refactoring, seems to lie around a token volume of 63 tokens. There are fewer large clones than small clones, resulting in a very limited statistical significance on our corpus when considering clones larger than 100 tokens.

In Table~\ref{tab:relation_refactor} we see the results regarding refactorings that are applied to clones with diverse relations. We see that most refactored clones are in a common class, over 54\%. This is significantly more than the percentage of clones in the common class relation as reported in Table~\ref{tab:relation}. Meanwhile, the number of refactored unrelated clones is smaller than the number reported in Table~\ref{tab:relation} (24\% -> 18\%). The main reason for this is that refactoring unrelated clones can change the relation of other clones in the same system. If we create a superclass abstraction to refactor an unrelated clone, other clones in those classes that were previously unrelated might become related.

The maintainability scores displayed in Table~\ref{tab:relation_refactor} show that the most favorable clones to refactor are clones with a Sibling relation. The most unfavorable is to refactor clones to interfaces. However, the differences in maintainability in this table are generally small; according to our data relations have only a minor impact on the maintainability of clones.

Regarding the return type of refactored clones, we see in Table~\ref{tab:return} that this has no major impact on maintainability. A method call to the extracted method that is directly returned and no return type extracted methods are slightly more favorable than the others. We think the main reason that the ``Return'' category is on top is that when a variable is declared at the end of the cloned fragment, CR directly returns its value and removes the declaration. This decreases the volume slightly.

A higher number of parameters directly influences the corresponding metric. Because of this, we see in Fig.~\ref{fig:arguments} that more parameters negatively influence maintainability. Not only the number of parameters metric is negatively influenced, but more method parameters also increase volume for the extracted method and each of the calls to it. Because of that, we see that the trend of the graph in Fig.~\ref{fig:arguments} decreases relatively rapidly.

\section{Conclusion} \label{sec:conclusion}
We defined automatically refactorable clones and created a tool to detect and refactor them. We measured statistical data with this tool over a large corpus of open-source Java software systems to get more information about the context of clones and how refactoring them influences system maintainability.

@@

We built a tool that can automatically apply refactorings to 28\% of the clones in our corpus using the ``Extract Method'' refactoring technique. The main reason our tool could not refactor all clones is that many clones span certain statements that obstruct method extraction, e.g., when code outside a method is part of a clone.

We measured four maintainability metrics before- and after applying each refactoring to determine the impact of each refactoring on system maintainability. We found that the most prominent factor influencing maintainability is the size of the clone. We found that the threshold lies at a clone volume of 29 tokens per clone instance for system maintainability to increase after refactoring the clone. Another factor with a major impact on maintainability is the number of parameters that the extracted method requires to get all required data. We noticed that the inheritance relation of the clone and the return value of the extracted method has only a minor impact on system maintainability.

>>>>

%Clone detection tools often use thresholds to determine which clones should be considered. Most clone detection tools use the number of lines \cite{kamiya2002ccfinder, svajlenko2016bigcloneeval}, statements \cite{higo2013revisiting, kamalpriya2017enhancing} and/or tokens \cite{ragkhitwetsagul2019siamese, jiang2007deckard} as thresholds to determine whether clones should be considered. We outline the shortcomings of these thresholds to find clones such that the maintainability of the source code is improved when refactored, and propose new thresholds to control this characteristic.
% Current clone detection techniques are limited to measuring the size of clones to determine whether they should be considered. Because of this, the output of clone detection tools is often of limited assistance to the developer in the refactoring process.
To reduce duplication of a codebase, there are several refactoring techniques that can be applied. The most used refactoring technique to deal with code clones is ``Extract Method'' \cite{fowler2018refactoring}. This technique entails that the duplicated code is moved to a new method, which is then called from all locations where the duplicated code was used. To be able to execute such a refactoring, several preconditions apply \cite{tsantalis2015assessing}.
\todo[inline]{still missing link on how current solutions do not help improving maintainability - our message from monday night is not crystal clear yet}
In this study, we analyze which preconditions apply to determine whether a clone could and should be automatically refactored. We list criteria that influence the maintainability of clones when refactored. We evaluate the context of clones to determine which refactoring techniques are best suited. We define the context of clones as the inheritance relation between clone instances in a clone class and the location of clone instances in the codebase. We showcase the proposed technique as a tool that automatically refactors clones by extracting new methods out of duplicated code.

@@

\section{Background}\label{sec:background}
We briefly describe relevant code clone and clone refactoring research.

@@

\subsubsection{Refactorability} \label{sec:refactorability}
Several factors may obstruct the possibility to extract code to a new method~\cite{tsantalis2015assessing}:
\begin{itemize}
    \item \textbf{Complex Control Flow}: This clone contains \texttt{break}, \texttt{continue} or \texttt{return} statements, obstructing the possibility of method extraction.
    \item \textbf{Spans Part Of A Block}: This clone spans a part of a block statement. An example of this is shown in Fig.~\ref{fig:refactorabilitypartialblock}.
    \item \textbf{Is Not A Partial Method}: If the clone does not fall in the ``Partial method'' category of Sec.~\ref{sec:location}, the ``Extract Method'' refactoring technique cannot be applied.
    \item \textbf{Returns Multiple Values}: If a clone modifies or declares multiple pieces of data that it should return.
    \item \textbf{Top-Level Non-Statement}: If one of the top-level AST nodes of the clone is not a statement. For instance, if a (part of) an anonymous class is cloned.
    \item \textbf{Can Be Extracted}: This clone is a fragment of code that can directly be extracted to a new method. Then, based on the relation between the clone instances, further refactoring techniques can be used to refactor the extracted methods (for instance \todo{be consistent: ``Extract Method" or ``extract method"} ``pull up method'' for clones in sibling classes).
\end{itemize}
Clones that do not fall in the \textit{Can Be Extracted} category may require additional transformations or other techniques to refactor. CR only refactors the clones that \textit{Can Be Extracted}.

\subsubsection{AST Transformation} % TODO: Big todo --> Find literature, introduce proper reasoning why I need to do this.
CR uses JavaParser~\cite{smith20e17javaparser}: an AST-parsing library that allows to modify the AST and write it back to source code. To refactor clones, CR creates a new method declaration and moves all statements from a clone instance in the clone class to the new method. This method is placed according to the relation between the clone instances (see Sec.~\ref{sec:relation}). CR analyzes the source code of the extracted method and populates it with the following properties:
\begin{itemize}
  \item \textbf{Parameters}: For each variable used that is not accessible from the scope of the extracted method.
  \item \textbf{A return value}: If the method modifies or declares local data that is needed outside of its scope, or if the cloned fragments already returned data.
  \item \textbf{Thrown exception}: If the method throws an uncaught exception that is not a \texttt{RuntimeException}.
\end{itemize}
CR then removes all cloned code and replaces it with a call to the newly created method. In case of a return value, CR either assigns the call result, declares it or returns it accordingly.

\subsubsection{Characteristics of the extracted method}\label{sec:characteristics}
We define the following characteristics of the extracted method and/or the call: \todo{why these?}
\begin{itemize}
\item \textbf{Tokens}: The number of tokens in the body of the method.
\item \textbf{Relation}: The relation category (Sec.~\ref{sec:relation}) by which this methods' location was determined.
\item \textbf{Returns}: Whether the method calls return, declare, assign or don't use any data from the extracted method.
\item \textbf{Parameters}: The number of parameters the extracted method has.
\end{itemize}
We hypothesize that these characteristics are the main factors influencing the impact on the maintainability of the system as a result of refactoring the clone. \todo{why?}

\subsubsection{Impact on maintainability metrics} \label{sec:metrics}
\todo[inline]{rephrase this text for clarity: this paper -- which paper? }
CR measures the impact on maintainability metrics of refactoring source code for each clone class that is refactored. These metrics are derived \todo{how did you change them?} from Heitlager et al.~\cite{heitlager2007practical}. This paper defines a set of metrics to measure the maintainability of a system. For each of these metrics, risk profiles are proposed to determine the maintainability impact on the system of a whole.

To determine whether the maintainability improves when refactoring a given clone, we need to measure the impact of fine-grained changes. Therefore, we measure only a subset of the metrics~\cite{heitlager2007practical} and focus on the absolute metric changes (instead of the risk profiles). The subset of metrics we choose to focus on are all measured on method level (as the other metrics show a lesser impact on the maintainability for these small changes). These metrics are:
\begin{itemize}
\item \textbf{Duplication}: Originally~\cite{heitlager2007practical} measured by taking the amount of duplicated lines. We decided to use the amount of duplicated tokens part of a clone class instead, to have a stronger reflection of the impact of the refactoring by measuring a more fine-grained system property.
%\item \textbf{Volume}: The more code a system has, the more code has to be maintained. The paper~\cite{heitlager2007practical} measures volume as lines of code. As with duplication, we use the number of tokens instead.
\item \textbf{Method Complexity}:  Originally ~\cite{heitlager2007practical} computed using MCCabe complexity~\cite{mccabe1976complexity}. The MCCabe complexity is a quantitative measure of the number of linearly independent paths through a method.\todo{why is the original way ok for us?}
\item \textbf{Method Interface Size}: The number of parameters that a method has. If a method has many parameters, the code may become harder to understand and it is an indication that data is not encapsulated adequately in objects~\cite{fowler2018refactoring}.
\item \textbf{Method size}: The longer a method is, the harder it becomes to comprehend and maintain \cite{heitlager2007practical}. The ``extract method'' refactoring technique is often used to split up long methods \cite{charalampidou2018structural}. When refactoring duplicated code fragments in the body of a method, this could improve maintainability of the refactored (reduced size).
\end{itemize}

\section{Experimental setup} \label{sec:experimentalsetup}
\todo{This doesn't say anything: explain how the section is organized}
%In this section, we describe the setup of our experiments.
Our goal is analysing whether automated refactoring of code clones leads to an improved maintainability. We first validate our automated refactoring tool, CR. Next, we run our tool on a reference corpus and measure the change in the maintainability score.

\subsection{Tool Evaluation}
We assess the correctness of CR through unit tests and empirical validation. First, we create a set of 57 control projects to verify the correctness in many (edge) cases. These projects contain clones for each relation, location, and refactorability category. 
%to see whether they get correctly identified. 
Next, we run the tool over the corpus and manually verify samples \todo{how many? why not all?} of the acquired results. This way, we check both the correctness of the identified clones, their context, and their proposed refactoring.

We also test the correctness of the refactored code using the JFreeChart project~\cite{jfreechart}. JFreeChart has a high test coverage and working tests, which allows us to test the correctness of the program after running CR.

\subsection{Corpus}
Our corpus, consisting of open source Java projects, is derived from the corpus of a study that uses machine learning to determine the suitability of GitHub Java projects for data analysis purposes~\cite{githubCorpus2013}.
\todo[inline]{why did you choose this corpus?}

CR requires all libraries of the analyzed projects
%it analyses, to find the full paths of all referenced symbols in the source code 
(see Sec.~\ref{sec:t1r}). Therefore, we decided to filter the corpus to contain only projects using the Maven build automation system. The filtering scripts are publicly availble \footnotesize{The camera-ready version would include the GitHub repo url}.
%We created a set of scripts to prepare such a corpus with all dependencies included.

This procedure results in 2.267 Java projects including all their dependencies \todo{this is confusing, how many original projects are there? were the dependencies in the original corpus?}. The projects vary in size and quality. The total size of all projects is 14.2M lines (11.3M when excluding whitespace) over a total of 100K Java files. This is an average of 6.3K lines over 44 files per project. The largest project in the corpus is \textit{VisAD} with 502K lines.
\todo[inline]{figure with the distribution of project sizes?}

\subsection{Minimum clone size}
% In this study, we want to find out what thresholds to use to improve maintainability if clones by those thresholds are refactored. However, 
When clones are very small, they may never be able to improve maintainability. The detrimental effect of the added volume of the newly created method exceeds the positive effect of removing duplication. We perform all our experiments with a minimum clone size of 10 tokens, as smaller clones are very unlikely\todo{any ref?} to improve maintainability when refactored.

\subsection{Thresholds}
Most clone detection tools can be configured using thresholds. These thresholds indicate the minimum number of lines, tokens and/or statements that must be spanned for duplicate fragments to be considered clones. Often, such thresholds are intuitively chosen~\cite{li2006cp, roy2009mutation} or based on a quartile \todo{quartile or quantile?} distribution of empirical data~\cite{alves2010deriving}. 
\todo{paragraph commented below seems more suitable for the discussion part. otherwise it should be rephrased.}
% Using the maintainability score we can find support for which thresholds should be chosen to increase the chance to find clones that improve maintainability when refactored.

@@

\subsection{Calculating a maintainability score}\label{sec:metricformula}
We use four metrics to determine maintainability (see Sec.~\ref{sec:metrics}). We determine the value of each metric before and after for each refactored clone class, resulting in a delta metric score. We aggregate the deltas obtained for these metrics to draw a conclusion about the maintainability increase or decrease after applying a refactoring. We base our aggregation on the following assumptions derived from supporting evidence~\cite{heitlager2007practical},~\cite{alves2010deriving}:
\begin{itemize}
  \item All metrics are equal in terms of weight towards system maintenance effort.
  \item Higher values for the metrics imply lower maintainability.
  \item Normalizing each metric delta over all deltas obtained for that metric in our dataset results in equally weighted
  scores.\todo{confusing}
  \item \todo{isn't this also an assumption?} all considered metrics do not interfere with the out-of-scope metrics
\end{itemize}

We use the resulting aggregated maintainability score to analyze  for each refactoring whether it increases or decreases the maintainability of the system.

We normalize each obtained metric delta using the ``Standard score''\todo{this name is confusing, is it established? then cite ref, otherwise I don't understand the quotes}, which is calculated as follows \todo{first give the intuition, then the math!}:
\begin{equation}\label{eq:scoredev}
N_{metric} = \frac {\Delta X-\mu}{\sigma}
\end{equation}
Where $\Delta X$ is a metric delta, $\mu$ is the mean of all deltas for this metric and $\sigma$ is the standard deviation of all deltas for this metric. This method works well for normalization of our data because as we divide by the standard deviation, outliers do not influence the resulting scores negatively \todo{backup by citing relevant reference}.

We calculate the maintainability of a refactoring as:
\begin{equation}\label{eq:scoreref}
N_{duplication} + N_{complexity} + N_{volume} + N_{parameters}
\end{equation}

@@

\section{Results} \label{sec:results}
We first report the context results, then the refactorability results.

@@

\subsubsection{Relation}
Table~\ref{tab:relation} shows the number of clone classes found for the entire corpus for each type of relations between clone instances (see Sec.~\ref{sec:relation}).

@@

\subsubsection{Location}
Table~\ref{tab:location} shows the number of clone classes found for the entire corpus for different locations (see Sec.~\ref{sec:location}).

@@

The results show that 74\% of clones span part of a method body (77\% if we include constructors). 8\% of clones span several methods. 6\% of clones span only global variables. Only 4\% of clones span a full declaration (method, class, constructor, etc.).

@@

\todo[inline]{the paragraph below belongs to discussion}
The results indicate that, given our refactorability criteria, 28\% of clones can be automatically refactored. Clones in other categories may require other refactoring techniques or further transformations to be automatically refactorable.

\subsection{Thresholds}
 CR has refactored 12.710 clone classes and measured the change in the selected metrics (Sec.~\ref{sec:metrics}). \todo{use the formulas labels to refer here} Using the presented formulas (Sec.~\ref{sec:metricformula}) we determine how the characteristics of the extracted method (see Sec.~\ref{sec:characteristics}) influence the maintainability of the resulting codebase after refactoring. We explore the data obtained by comparing the before- and after snapshots of the system for each separate refactoring. Using this data, we find supporting evidence regarding which thresholds are most likely to find clones that should be refactored to improve maintainability.

\subsubsection{Clone Token Volume}
Figure \ref{fig:maintainabilityscore} shows the results by plotting the clone volume vs the average delta maintainability score. We define the \textit{token volume} as the combined number of tokens in all clone instances in a refactored clone class. For higher token volume numbers we have fewer refactorings that refactor such clones, therefore we represent the x-axis as a logarithmic scale. The trendline intersects the ``zero'' line (maintainability does not increase nor decrease) at a token volume of 63.

@@

\subsubsection{Relation}
Table~\ref{tab:relation_refactor} shows our data regarding how different relations influence maintainability. We have marked rows based on less than 100 refactorings red\todo{not color-agnostic, if printed bw might be a problem --> just make it grey }, as their result does not have statistical significance. Relations are ordered by their scores. \todo{this sentence about relation impact seems more suitable for discussion}Scores do not deviate much (-0.15 to 0.23), indicating that the relation between clones has a minor impact on maintainability. Overall, common hierarchy clones have the highest maintainability, whereas unrelated clones have the lowest maintainability.

\subsubsection{Return Value}
Table~\ref{tab:return} shows how the return value of the extracted method influences the maintainability of the resulting system. The return categories are ordered by their scores. \todo{this sentence about relation impact seems more suitable for discussion}Scores do not deviate much (-0.18 to 0.19), indicating that the return category has a minor impact on maintainability. When the result of the call to the extracted method is directly returned, the maintainability score is the highest. When no value is returned, the maintainability score is the lowest. \todo{this sentence about relation impact seems more suitable for discussion}The main reason that no return value is lowest, is that it is linked to a higher number of parameters required for the extracted method.

@@

@@

\subsection{Clone Context}
Regarding clone context, our results indicate that most clones (37\%) are in a common class. This is favorable for \todo{automated?} refactoring because the extracted method does not have to be moved after extraction. 24\% of clones are in a common hierarchy. These refactorings are also often favorable. Another 24\% of clones are unrelated, which is often unfavorable because it often requires a more comprehensive refactoring. 15\% of clones are in an interface \todo{which ...}.

@@

\subsection{Extract Method}
28\% of clones can be refactored using the ``Extract Method'' refactoring technique (50\% if we limit our searching scope to method bodies). About 25\% of clones do not span part of a method, therefore they cannot be refactored. Many clones (23\%) do not have a statement as top-level AST-Node. Upon manual inspection, we noticed that the main reason is clones in lambda functions or in anonymous classes. About 15\% of clones span only part of an AST-Node \todo{and?}.

\subsection{Refactoring}
Fig.~\ref{fig:maintainabilityscore} shows an increase in maintainability for refactoring larger clone classes. The tipping point, between a better and a worse maintainable refactoring, seems to lie around a token volume of 63 tokens. There are fewer large clones than small clones, resulting in a very limited \todo{how limited? give the p-value} statistical significance on our corpus when considering clones larger than 100 tokens.

Table~\ref{tab:relation_refactor} shows the results regarding refactorings that are applied to clones with diverse relations. More than 54\% of the refactored clones are in a common class. This is significantly more than the percentage of clones in the common class relation reported in Table~\ref{tab:relation}. The number of refactored unrelated clones is smaller than the number reported in Table~\ref{tab:relation} (24\% -> 18\%). The main reason for this is that refactoring unrelated clones can change the relation of other clones in the same system. If we create a superclass abstraction to refactor an unrelated clone, other clones in those classes that were previously unrelated might become related.

The maintainability scores in Table~\ref{tab:relation_refactor} show that the most favorable clones to refactor are clones with a Sibling relation. The most unfavorable is refactoring clones to interfaces. However, the differences in maintainability in this table are generally small \todo{and?}; according to our data relations have only a minor impact on the maintainability of clones.

Regarding the return type of refactored clones, Table~\ref{tab:return} shows that this has no major impact on maintainability. A method call to the extracted method that is directly returned and no return type extracted methods are slightly more favorable than the others. We think the main reason that the ``Return'' category is on top is that when a variable is declared at the end of the cloned fragment, CR directly returns its value and removes the declaration. This decreases the volume slightly.

% A higher number of parameters directly influences the corresponding metric. 
Fig.~\ref{fig:arguments} shows that more parameters negatively influence maintainability. Not only the \todo{wasn't this a characteristic? this paragraph is somewhat confusing} number of parameters metric is negatively influenced, but more method parameters also increase volume for the extracted method and each of the calls to it. Because of that, we see that the trend of the graph in Fig.~\ref{fig:arguments} decreases relatively rapidly.

\todo[inline]{one paragraph with a general remark on this discussion, summarizing our findings}
\section{Conclusion} \label{sec:conclusion}
\todo[inline]{one paragraph summarizing the context of our research}
We defined automatically refactorable clones and created a tool to detect and refactor them. We measured statistical data with this tool over a large corpus of open-source Java software systems to get more information about the context of clones and how refactoring them influences system maintainability.

@@

We built a tool that can automatically apply refactorings to 28\% of the clones in our corpus using the ``Extract Method'' refactoring technique. 
% The main reason our tool could not refactor all clones is that many clones span certain statements that obstruct method extraction, e.g., when code outside a method is part of a clone.

We measured the change in four maintainability metrics
%before- and after applying each refactoring 
to determine the impact of each refactoring on system maintainability. We found that the most prominent factor influencing maintainability is the size of the clone. We found that the threshold lies at a clone volume of 29 \todo{wasn't this 63?} tokens per clone instance for system maintainability to increase after refactoring the clone. Another factor with a major impact on maintainability is the number of parameters that the extracted method requires to get all required data. We noticed that the inheritance relation of the clone and the return value of the extracted method has only a minor impact on system maintainability.
